# Matrices

We use the `matrix` function to create a matrix from a vector of values.

```{r mat_define}
A <- matrix(c(1,2,4,8), nrow=2, ncol=2, byrow=FALSE)
A
B <- matrix(c(1,2,4,8), nrow=2, ncol=2, byrow=TRUE)
B
```

We can extract elements of the matrix with square brackets.
Look up the documentation with `?"["`.
We can extract elements by row
```{r mat_index_row}
A[1,]
A[2,]
```

by column

```{r mat_index_col}
A[,1]
A[,2]
```

or by element

```{r mat_index_elem}
A[1,2]
A[2,1]
```

We can also use square brackets to replace columns, rows, or elements of a matrix.

```{r mat_replace}
A[1,] <- c(1, 2)
A
A[,2] <- c(2, 1)
A
A[2,1] <- -1
A
```

The arithmetic operators perform elementwise operations for matrices, as with vectors.
```{r mat_arithmetic}
A + B
A - B
A * B
A / B
A ^ B
```

To perform matrix multiplication, we use the `%*%` operator.
```{r mat_mult}
A %*% B
```

To find the inverse of a matrix, we use the `solve` function.
```{r mat_inverse}
solve(A)
```

We also use the `solve` function to solve the equation $Ax = b$.
```{r mat_solve}
solve(A, b=c(3,9))
```

To find the determinant of a matrix, we use the `det` function.
```{r mat_determinant}
det(A)
```

We can write functions to extract a general matrix diagonal
```{r mat_gendiag}
gen.diag <- function(A, k) A[row(A) == col(A) - k]
```

and to replace a general matrix diagonal
```{r mat_gendiagassg}
`gen.diag<-` <- function(A, k, value) {
  A[row(A) == col(A) - k] <- value
  A
}
```

Let's test these functions for extracting and replacing a general matrix diagonal.
```{r mat_gendiag_test}
C <- matrix(c(1,0,-1,8,1,0,-1,9,-2), nrow=3, byrow=TRUE)
C
gen.diag(C, -1)
gen.diag(C, -1) <- c(2,1)
C
```
